---
title: "CoinGecko integration: pulling crypto data without an API key"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CoinGecko integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, eval = FALSE, purl = FALSE, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE,    # Don't run at vignette-build time
  purl     = FALSE     # Don't include in tangled .R either - R CMD check
                       # runs the tangled script and would otherwise try
                       # to install remotes / hit the live CoinGecko API.
)
```

# Why a CoinGecko branch?

`crypto2` was built around CoinMarketCap (CMC) because CMC was historically the
only free source of **survivorship-bias-free historical crypto listings** —
their internal API serves delisted-coin data without an API key, which is
exactly what academic finance research needs.

CMC's internal API is also notoriously unstable — the `crypto_info()` function
has been patched several times because CMC keeps adding fields that break the
deny-list approach. The CoinGecko branch is the second source: a CG-side
counterpart with the same column conventions, so downstream research code that
already consumes a `crypto_list()` / `crypto_listings()` / `crypto_history()`
tibble works on a `cg_list()` / `cg_listings()` / `cg_history()` tibble too.

Triangulating across both sources is the safest way to insulate your research
against any one platform changing its policies.

The branch is experimental. Install with:

```{r, eval = FALSE, purl = FALSE}
# install.packages("remotes")
remotes::install_github("sstoeckl/crypto2@coingecko-integration")
library(crypto2)
```

# The endpoint surface

CoinGecko exposes two different hosts:

| Host | Purpose | Rate limit | Key required? |
|---|---|---|---|
| `api.coingecko.com/api/v3/...` | Documented public API. Demo tier, stable schema, versioned. | **30 req / min** (Demo, key-less) | No |
| `www.coingecko.com/...` | Undocumented internal endpoints that the CoinGecko **website itself** uses to render its pages. JSON responses, browser-like User-Agent recommended. | No documented cap; Cloudflare bot filtering at high request rates | No |

Inside this package, every `cg_*` function picks the host that makes most
sense for its job. The summary:

| Function | Hits | Rate-limit posture |
|---|---|---|
| `cg_list()` | documented `coins/list` + `coins/markets` paginated | Sleeps **2.5 s** between page calls (≈ 24/min, 80 % of cap) |
| `cg_listings()` | documented `coins/markets` paginated | Same |
| `cg_history()` | undocumented `price_charts/`, `market_cap/`, `ohlc/` per coin | Sleeps **0.6 s** (≈ 100/min, polite) |
| `cg_history_by_id()` | same as `cg_history()` but addresses by numeric ID | Same |
| `cg_info()` | documented `coins/{slug}` per coin | Sleeps 2.5 s |

# The four core functions

All four return tibbles with column names that mirror their `crypto_*`
counterparts.

## `cg_list()` — the active coin universe

```{r, eval = FALSE, purl = FALSE}
universe <- cg_list(top_n = NULL)        # ≈ 17 000 coins (active only)
head(universe)
#> # A tibble: 6 × 8
#>      id name       symbol slug      rank is_active first_historical_data last_historical_data
#>   <int> <chr>      <chr>  <chr>    <int>     <int> <date>                <date>
#> 1     1 Bitcoin    btc    bitcoin      1         1 NA                    2026-05-13
#> 2   279 Ethereum   eth    ethereum     2         1 NA                    2026-05-13
#> 3   325 Tether     usdt   tether       3         1 NA                    2026-05-13
#> ...
```

* `id`: CoinGecko **internal numeric ID** (extracted from the image URL in
  `/coins/markets`). The numeric ID is the primary key for the
  website-host endpoints (`/price_charts/{id}/...` etc.).
* `slug`: CoinGecko **API slug** (= the `id` field in CG's documented API).
  Used for the slug-addressed endpoints.
* `rank`: current market-cap rank (`market_cap_rank` from CG).
* `is_active`: always `1L` on CG's free tier (CG does not expose inactive
  coins via `/coins/list` without a paid key — see the
  *survivorship-bias-free strategy* section below).

`top_n` controls how many pages of `/coins/markets` are pulled for
enrichment. `top_n = 0` skips enrichment entirely (one HTTP call, no
numeric IDs); `top_n = NULL` paginates the full universe.

## `cg_listings()` — current cross-sectional snapshot

```{r, eval = FALSE, purl = FALSE}
snapshot <- cg_listings(limit = 1000)     # top 1000 by market cap, with returns
head(snapshot)
```

Returns 30 columns that match `crypto_listings(quote = TRUE)` where the
corresponding CG field exists:

* `id, name, symbol, slug`
* `last_updated`, `rank`
* `market_cap, fully_diluted_market_cap`
* `circulating_supply, total_supply, max_supply`
* `price, volume_24h, high_24h, low_24h`
* `percent_change_1h, _24h, _7d, _14d, _30d, _200d, _1y`
* `ath, ath_change_percentage, ath_date`, same for `atl`
* `ref_currency` (the upper-cased quote currency, matching `crypto_listings()`'s
  CMC output)

Historical cross-sectional snapshots (`which = "historical"`) are **not
available** on CG's free tier — see the next section.

## `cg_history()` — daily OHLC + volume + market cap per coin

```{r, eval = FALSE, purl = FALSE}
# Pass a tibble in cg_list()/cg_listings() format
top10 <- cg_list(top_n = 10)
hist  <- cg_history(coin_list = top10,
                    what       = c("price", "market_cap", "ohlc"),
                    start_date = "2020-01-01")
```

The output schema matches `crypto_history()`:

```
id, slug, name, symbol, timestamp,
ref_cur_id, ref_cur_name,
open, high, low, close, volume, market_cap,
time_open, time_high, time_low, time_close
```

Three calls per coin per stream — one each to `/price_charts/`,
`/market_cap/`, `/ohlc/` (the OHLC call requires the numeric `id` from
`cg_list()`). Each call returns the coin's **full** history in one
response; the `start_date` / `end_date` filtering happens client-side.

OHLC sparsity caveat: CoinGecko returns daily OHLC candles only for
windows ≥ 91 days back from "now"; recent periods return finer (4-hour)
candles that don't align to UTC midnight, so the daily floor drops them
from the OHLC join. Price / Volume / Market-cap remain perfectly daily.

## `cg_info()` — per-coin metadata

```{r, eval = FALSE, purl = FALSE}
info <- cg_info(coin_list = data.frame(slug = "bitcoin"))
```

Schema mirrors `crypto_info()` where there is a direct equivalent, plus a
few CG-specific extras (`platforms` for cross-chain contract addresses,
`categories` for thematic groupings).

# The survivorship-bias problem

CoinGecko **deliberately prunes** delisted coins from its public API.
From the platform's own support documentation:

> Users will not be able to access historical data for inactive or
> delisted coins via the CoinGecko API.

> CoinGecko periodically reviews existing listings and if they fail to
> meet their requirements, then they will be deactivated or removed.

> CoinGecko has removed over 24 000 "dead" cryptocurrencies from its
> platform.

For academic finance research this is disqualifying: any cross-section
you assemble from CG today is the **surviving subset**, which biases
returns upward, volatility downward, and breaks every cross-sectional
factor-pricing test.

There are **three** paths to a survivorship-bias-free dataset on
CoinGecko, with different trade-offs:

# Strategy A — Accumulate via a weekly cronjob (free, slow ramp-up)

The simplest path. Every week, snapshot `cg_list()` (active universe at
that point) and persist it. Coins that get delisted **after** you start
snapshotting remain in your archive. The union of all snapshots grows
into a survivorship-bias-corrected universe over time.

```{r, eval = FALSE, purl = FALSE}
# inside your cronjob (separate package):
weekly_universe <- cg_list(top_n = NULL)
arrow::write_parquet(
  weekly_universe,
  file.path("data/cg_universes", paste0(Sys.Date(), ".parquet"))
)
```

After a few months you have a usable rolling bias-corrected universe.

# Strategy B — One-shot Pro-tier harvest, then free forever ($129)

The **best** strategy if you want a complete *historical* universe right
now, including coins that were delisted before you started.

Key insight: CoinGecko's internal numeric IDs are stable and the
**website-host endpoints (`/price_charts/{id}/...`, `/market_cap/{id}/...`,
`/ohlc/{id}/series/...`) continue to serve historical data even for
delisted coins, addressed by numeric ID**. The slug routing is what gets
removed when CG delists a coin, not the underlying numeric-ID data.

So: pay for one month of the **Analyst tier** ($129), harvest the
complete (slug, numeric_id, name, symbol, status) mapping for the entire
universe (active **and** inactive), then cancel the subscription. From
then on, you use the free `cg_history_by_id()` against your harvested
numeric IDs.

The Pro tier gives you:

* `pro-api.coingecko.com/api/v3/coins/list?status=inactive` — the
  delisted-coin slug list (not available on the Demo tier).
* `pro-api.coingecko.com/api/v3/coins/{slug}` for **delisted** slugs —
  returns the image URL (and thus the numeric ID), plus name, symbol,
  genesis date.

## Step 1 — get the API key

Subscribe to the [Analyst tier](https://www.coingecko.com/en/api/pricing).
Copy your API key. Set it as an environment variable:

```{r, eval = FALSE, purl = FALSE}
Sys.setenv(CG_PRO_KEY = "your-key-here")
```

## Step 2 — harvest the universe master (`pro-api.coingecko.com`)

The package does not ship a Pro-tier harvest function (those endpoints
require a paid key, which is out of scope for the free package). Use
this self-contained script (adapt as needed):

```{r, eval = FALSE, purl = FALSE}
# Run once with an Analyst-tier key. Writes cg_universe_master.parquet
# with all the slug ↔ numeric_id ↔ name ↔ symbol mappings.
library(httr)
library(jsonlite)
library(purrr)
library(tibble)
library(dplyr)
library(stringr)
library(arrow)

key <- Sys.getenv("CG_PRO_KEY")
stopifnot(nzchar(key))

pro_get <- function(path, query = NULL) {
  url  <- sprintf("https://pro-api.coingecko.com/api/v3/%s", path)
  resp <- httr::GET(url, query = query,
                    httr::add_headers(`x-cg-pro-api-key` = key,
                                      `User-Agent` = "crypto2/coingecko-harvest"),
                    httr::timeout(60))
  if (httr::status_code(resp) != 200) return(NULL)
  jsonlite::fromJSON(httr::content(resp, "text", encoding = "UTF-8"))
}

# 1. Active + inactive universes (slug lists)
active   <- pro_get("coins/list",
                    query = list(include_platform = "true"))
inactive <- pro_get("coins/list",
                    query = list(include_platform = "true",
                                 status           = "inactive"))

all_slugs <- unique(c(active$id, inactive$id))
status_lookup <- setNames(
  c(rep("active",   length(active$id)),
    rep("inactive", length(inactive$id))),
  c(active$id, inactive$id)
)

# 2. Per-slug detail → numeric_id, name, symbol, genesis_date
extract_one <- function(slug) {
  d <- pro_get(sprintf("coins/%s", slug),
               query = list(localization   = "false",
                            tickers        = "false",
                            market_data    = "false",
                            community_data = "false",
                            developer_data = "false",
                            sparkline      = "false"))
  if (is.null(d)) return(NULL)
  image_url  <- d$image$large %||% d$image$small %||% d$image$thumb
  numeric_id <- as.integer(stringr::str_extract(
    image_url, "(?<=/coins/images/)\\d+"))
  tibble::tibble(
    numeric_id        = numeric_id,
    slug              = slug,
    name              = d$name,
    symbol            = d$symbol,
    status            = status_lookup[[slug]],
    genesis_date      = as.Date(d$genesis_date %||% NA_character_),
    asset_platform_id = d$asset_platform_id %||% NA_character_,
    web_slug          = d$web_slug %||% slug,
    harvested_at      = Sys.time()
  )
}

# At ~500 req/min on the Analyst tier, ~40 000 calls = ~80 min.
mapping <- purrr::map_dfr(all_slugs,
                          purrr::possibly(extract_one, otherwise = NULL),
                          .progress = TRUE)

arrow::write_parquet(mapping, "cg_universe_master.parquet")
```

This `cg_universe_master.parquet` is the **Rosetta stone**. After this
finishes, cancel your Analyst subscription.

## Step 3 — use the free `cg_history_by_id()` going forward

```{r, eval = FALSE, purl = FALSE}
mapping <- arrow::read_parquet("cg_universe_master.parquet")

# Refresh history for the entire historical universe — free, no key.
# Slow (~12 h at default sleep) but only needed once for the backfill.
hist <- cg_history_by_id(
  ids       = mapping$numeric_id,
  what      = c("price", "market_cap", "ohlc"),
  coin_list = mapping       # for slug join on rows where mapping has it
)
arrow::write_parquet(hist, "cg_history_backfill.parquet")
```

## Step 4 — keep the master mapping fresh going forward (free)

Run weekly: pull `cg_list()` (free), append any **new** (slug,
numeric_id) pairs to the master mapping. The mapping grows organically
as CoinGecko lists new coins.

```{r, eval = FALSE, purl = FALSE}
weekly <- cg_list(top_n = NULL)
new_pairs <- weekly %>%
  filter(!is.na(id), !id %in% mapping$numeric_id) %>%
  transmute(numeric_id = id, slug, name, symbol,
            status = "active", genesis_date = NA,
            harvested_at = Sys.time())

mapping <- bind_rows(mapping, new_pairs)
arrow::write_parquet(mapping, "cg_universe_master.parquet")
```

Future delistings: when a coin is removed from `/coins/list`, its slug
disappears from your weekly captures — but you've already preserved its
numeric ID in `mapping`, so `cg_history_by_id()` can still refetch its
history.

## What the one-shot strategy gives you

* **Backfill**: all coins ever listed on CoinGecko, including the
  pre-existing delisted ones (which weekly accumulation alone cannot
  recover).
* **Forward bias correction**: future delistings are automatically
  preserved via numeric-ID retention.
* **Cost**: $129 once. No recurring fee.
* **Limitation**: coins delisted *before* CoinGecko started tracking
  them at all are obviously not recoverable.

# Strategy C — cross-reference with CoinMarketCap (free)

`crypto2`'s existing `crypto_list(only_active = FALSE, add_untracked = TRUE)`
returns the *entire* CMC universe, including delisted coins. CMC's
delisting policy is stricter than CG's, so cross-referencing both sources
gives you:

* A second independent universe estimate.
* Coverage for coins that one platform delisted but the other kept.
* Cross-validation of market-cap / volume figures (within the limits of
  exchange aggregation differences).

```{r, eval = FALSE, purl = FALSE}
cmc_full <- crypto_list(only_active = FALSE, add_untracked = TRUE)
cg_full  <- arrow::read_parquet("cg_universe_master.parquet")

# Match on symbol+name (slugs differ across platforms)
joint <- dplyr::inner_join(cmc_full, cg_full,
                           by = c("symbol", "name"),
                           suffix = c(".cmc", ".cg"))
```

# Performance & rate-limit reference

| Function | Calls per run (typical) | Default sleep | Throughput | Notes |
|---|---:|---:|---|---|
| `cg_list(top_n = NULL)` | 1 + ~70 | 2.5 s | ~2 min | docu host, 24/min |
| `cg_listings()` | ~70 | 2.5 s | ~2 min | docu host |
| `cg_history(top_N coins)` | 3 × N | 0.6 s | ~30 coins/min | website host |
| `cg_history_by_id(union)` | 3 × N | 0.6 s | ~30 coins/min | website host |
| `cg_info(N coins)` | N | 2.5 s | 24/min | docu host |

For the full 40 000-coin backfill via `cg_history_by_id()`, expect
roughly 12 hours per stream (`price` / `market_cap` / `ohlc`), totaling
~36 hours wall-clock. Run as a background process or split into chunks.

# FAQ

**Q: Do I have to use the Pro-tier strategy?**
No. Strategy A (weekly cronjob) is fully free and builds the
bias-corrected universe over time. Strategy B is the fastest path to a
*backfilled* dataset.

**Q: What if CoinGecko closes the undocumented endpoints?**
The package's `cg_history_by_id()` would stop working. Your already-
harvested data remains intact. You'd need to migrate to a paid tier or
another data source for fresh history. This risk is inherent in any
undocumented-API approach, including `crypto2`'s use of CMC's internal
endpoints.

**Q: Are CoinGecko's numeric IDs sequential?**
Only in the **dense early range** (low IDs 1–~50 000 are mostly
populated). The full ID space extends to ~10⁸ but is **sparse** above
~70 000 — modern listings get pseudo-random IDs in that large space.
Blind `1:N` iteration does **not** recover the full universe; that's
why `cg_history_by_id()` defaults to using `cg_list()$id`, not a
numeric range.

**Q: Where's the cronjob package?**
The cronjob orchestration lives in a separate project — this package
ships only the data-pulling functions.
