---
title: "CoinGecko Pro backfill: one-shot survivorship-bias archive"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CoinGecko Pro backfill}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, eval = FALSE, purl = FALSE, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE,
  purl     = FALSE
)
```

# Scope

The free-tier `cg_*` functions are the right primary entry point for almost
all `crypto2` users — they require no key and accept the same arguments as
their CMC counterparts.

For one specific scenario — bootstrapping a **survivorship-bias-corrected
archive from scratch in a single batch run** — the Pro tier
(`pro-api.coingecko.com`) is the cheapest path: a one-shot subscription gets
you per-coin OHLC and listing snapshots for every coin CoinGecko has ever
tracked, in a few hours rather than the months of accumulated snapshots that
the free tier requires.

This vignette holds the **recipes**. The functions are written inline rather
than exported by `crypto2` — they are deliberately kept out of the package
namespace so that:

* there is **no encouragement** of paid-API patterns inside a key-free
  package, and
* the recipes can be adapted to any change in the Pro endpoints without
  bumping the package version.

To use the recipes: copy the function definitions below into a script,
provide your Pro API key, and run.

# Setup

```{r, eval = FALSE, purl = FALSE}
library(crypto2)         # only for the column conventions we mirror
library(dplyr)
library(tibble)
library(purrr)
library(jsonlite)
library(httr)
library(arrow)

# Your Pro key. Store in .Renviron as COINGECKO_PRO_KEY and read here.
CG_PRO_KEY <- Sys.getenv("COINGECKO_PRO_KEY", unset = NA)
stopifnot(!is.na(CG_PRO_KEY))
```

# A polite Pro client

```{r, eval = FALSE, purl = FALSE}
# Pro tier nominal cap: 500 req / min. Stay below ~ 6 req / s.
pro_sleep <- 0.2

pro_get <- function(path, query = NULL) {
  url <- paste0("https://pro-api.coingecko.com/api/v3/", sub("^/", "", path))
  resp <- httr::GET(
    url,
    query = query,
    httr::add_headers(`x-cg-pro-api-key` = CG_PRO_KEY),
    httr::timeout(60)
  )
  sc <- httr::status_code(resp)
  if (sc == 429) {
    ra <- suppressWarnings(as.numeric(
      httr::headers(resp)[["retry-after"]]))
    if (is.na(ra)) ra <- 30
    Sys.sleep(ra)
    return(pro_get(path, query))  # one retry on 429
  }
  if (sc < 200 || sc >= 300) return(NULL)
  jsonlite::fromJSON(httr::content(resp, as = "text", encoding = "UTF-8"))
}
```

# Recipe 1: full historic id/slug mapping

The Pro endpoint `/coins/list?include_platform=false&status=active,inactive`
returns *every* coin CoinGecko has ever tracked, active or not. This is the
input to the survivorship-bias-corrected universe.

```{r, eval = FALSE, purl = FALSE}
pro_id_mapping <- function() {
  raw <- pro_get("coins/list",
                 query = list(include_platform = "false",
                              status = "active,inactive"))
  if (is.null(raw) || !length(raw)) return(tibble::tibble())
  tibble::tibble(
    slug         = raw$id,
    symbol       = raw$symbol,
    name         = raw$name,
    harvested_at = Sys.Date()
  )
}

mapping <- pro_id_mapping()
nrow(mapping)
#> [1] ~ 17 000 (active) + 5 000-10 000 inactive
```

To enrich with the numeric CoinGecko IDs (needed for the OHLC website-host
endpoints), page through `/coins/markets`:

```{r, eval = FALSE, purl = FALSE}
pro_numeric_ids <- function() {
  per_page <- 250L
  pages <- vector("list", 200L)
  for (i in seq_along(pages)) {
    Sys.sleep(pro_sleep)
    page <- pro_get("coins/markets",
                    query = list(vs_currency = "usd",
                                 per_page = per_page, page = i))
    if (is.null(page) || !nrow(page)) break
    pages[[i]] <- tibble::tibble(
      slug   = page$id,
      id     = as.integer(sub("^.*/coins/images/([0-9]+).*$", "\\1",
                              page$image)),
      rank   = page$market_cap_rank
    )
    if (nrow(pages[[i]]) < per_page) break
  }
  dplyr::bind_rows(pages)
}

ids <- pro_numeric_ids()
mapping_full <- dplyr::left_join(mapping, ids, by = "slug")
```

# Recipe 2: full historic OHLC per coin

Pro `/coins/{slug}/ohlc?vs_currency=...&days=max` returns daily OHLC for
the entire history of the coin in a single call.

```{r, eval = FALSE, purl = FALSE}
pro_ohlc_one <- function(slug, vs = "usd") {
  Sys.sleep(pro_sleep)
  raw <- pro_get(sprintf("coins/%s/ohlc", slug),
                 query = list(vs_currency = vs, days = "max"))
  if (is.null(raw) || !length(raw)) return(NULL)
  tibble::tibble(
    slug      = slug,
    timestamp = as.POSIXct(raw[, 1] / 1000, origin = "1970-01-01", tz = "UTC"),
    open      = raw[, 2],
    high      = raw[, 3],
    low       = raw[, 4],
    close     = raw[, 5]
  )
}

# Run for the entire universe
hist <- purrr::map_dfr(mapping_full$slug, pro_ohlc_one)
```

# Recipe 3: persist as a parquet dataset

The accumulated parquet is the survivorship-bias-corrected archive. Combined
with the id mapping it lets `cg_history()` and `cg_list(only_active = FALSE)`
work correctly on the free tier forever after.

```{r, eval = FALSE, purl = FALSE}
arrow::write_parquet(mapping_full, "cg_id_mapping_pro.parquet")
arrow::write_dataset(
  hist,
  path        = "data/cg_history_pro",
  partitioning = "slug"
)
```

# Where to host the mapping for other users

If you intend the mapping to be reused by `cg_id_mapping()` (in this
package or by other consumers), upload the parquet to a stable, anonymous
public URL. The default download path baked into `cg_id_mapping()` is the
Hugging Face dataset
`sstoeckl/opencryptoassetpricing` at `data/_static.parquet`. Drop your
parquet there (after stripping anything but the four columns
`id`, `slug`, `symbol`, `name`, `harvested_at`) and the free-tier package
will pick it up automatically.

# Rate-limit budgeting

A one-shot historical bootstrap of ~ 17 000 coins at 0.2 s per call is
about **57 minutes** for the OHLC sweep alone, plus a few minutes for the
mapping and listing snapshot. Plan for ~ 2 hours total wall-clock with
generous safety margins. The Pro 30-day trial period is sufficient for
exactly one bootstrap.
